---
title: "Visualization"
output: github_document
date: "2026-10-01"
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")
```

now we have everything we need

```{r}
weather_df
```

let's make a scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

i always like dataframe first

```{r}
ggp_temp_scatterplot = 
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_temp_scatterplot
```

let's make this a bit fancier

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE)
```

alpha makes plots more transparent
geom_smooth shows best fit line
se = FALSE gets rid of margin of error

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = 0.25) +
  geom_smooth(se = FALSE)
```

the aesthetics are up to you

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

faceting

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(. ~ name)
```

faceting separates variables into different graphs

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  facet_grid(cols = vars(name))
```

let's look at something else

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) + 
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name)
```

make a plot of central park tmax v tmin only, convert temperatures to fahrenheit

```{r}
weather_df |>
  filter(name == "CentralPark_NY")|>
  mutate(
    tmax = tmax * (9/5) + 32,
    tmin = tmin * (9/5) + 32
  ) |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

filter to edit specific variable
mutate to create a formula

hex plot

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

geom_hex shows distribution

## univariate plots

```{r}
weather_df |>
  ggplot(aes(x = tmax)) +
  geom_histogram()
```

```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

fill colors in bars, good for histograms

density plots

```{r}
weather_df |>
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3)
```

easier to see the overlap between subvariables

boxplots

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

put categorical variable on x axis to separate

violin plots

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  geom_violin()
```

ridge plots

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```

ridge plot separates density plots into different plots...shows overlap still

## save some of my plots

```{r}
ggp_weather =
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  facet_grid(. ~ name)

ggp_weather

ggsave("images/ggplot_weather.pdf", ggp_weather)
```

```{r fig.width = 12, fig.height = 12}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

