---
title: "Visualization 2"
output: github_document
date: "2026-10-06"
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

data("weather_df")
```

now we have everything we need

let's make a scatterplot

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature (C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NDAA for three weather stations."
  )
```

let's try other scales

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature (C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NDAA for three weather stations."
  ) + 
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10 C,", "0", "Fifteen")
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

let's look at color

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  scale_color_hue(h = c(100, 300))
```

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```

## colors

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme(legend.position = "bottom")
```

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme_classic() +
  theme(legend.position = "bottom")
```

be aware of order of "theme"...sensitive

update tmax vs date plot

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    title = "Seasonal trends in Max Temp",
    x = "Date",
    y = "Max Temperature (C)",
    caption = "Max daily temp in three weather locations in 2021 and 2020",
    color = "Location"
  ) +
   viridis::scale_color_viridis(
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## two more weird but useful plot things

```{r}
central_park_df =
  weather_df |>
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |>
  filter(name == "Molokai_HI")

ggplot(molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

as long as the dfs has the same x and y variables, can use geom_line to show a different df

multiple panels with different plot types

```{r}
ggp_tmax_tmin =
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  theme(legend.position = "none")

ggp_prcp_density =
  weather_df |>
  filter(prcp > 0) |>
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

ggp_seasonal =
  weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prcp_density)/ggp_seasonal
```

plus sign indicates adding plots together
divide sign indicates putting plot on the bottom

## data manipulation

start with factors

boxplots

```{r}
weather_df |> 
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |>
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

bc weather_df is a character variable, ggplot automatically makes data into a factor (1, 2, 3)...here it is alphabetical order

relevel creates what order you want variables listed

```{r}
weather_df |> 
  mutate(name = fct_reorder(name, tmax)) |>
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

reorder name based on low to high 

make specific distribution plot

```{r}
weather_df |>
  select(name, date, tmax, tmin) |>
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temp"
  ) |> 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```

make FAS plot

```{r}
pups_df =
  read_csv("FAS_pups.csv", skip = 3, na = c("", ".", "NA")) |> janitor::clean_names() 

litters_df =
  read_csv("FAS_litters.csv", na = c("", ".", "NA")) |> janitor::clean_names() |> 
separate(group, into = c("dose", "day_of_tx"), 3)

fas_df =
  left_join(pups_df, litters_df, by = "litter_number")

fas_df = 
  fas_df |> 
  select(dose, day_of_tx, starts_with("pd")) |> 
  pivot_longer(
    cols = starts_with("pd"),
    names_to = "outcome",
    values_to = "pn_day"
  )

ggplot(fas_df, aes(x = dose, y = pn_day)) +
  geom_boxplot() + 
  facet_grid(day_of_tx ~ outcome)
```