---
title: "Numerical EDA"
output: github_document
date: "2026-10-08"
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

load weather data

```{r}
data("weather_df")

weather_df = 
  weather_df |> 
  mutate(
    month = lubridate::floor_date(date, unit = "month")
  )
```

now we have everything we need

```{r}
weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp)) +
  geom_histogram()
```

```{r}
weather_df |> 
  filter(prcp > 0)
```

```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x = tmax, y = tmin, color = name, shape = name)) +
  geom_point()
```

## `group_by`

```{r}
weather_df |> 
  group_by(name, month)
```

## `summerize()`

```{r}
weather_df |> 
  group_by(name) |> 
  summarize(count = n())
```

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(count = n())
```

```{r}
weather_df |> 
  group_by(month) |> 
  summarize(
    count = n(),
    n_days = n_distinct(date))
```

don't do this

```{r}
weather_df |> 
  pull(tmax) |> 
  summary()
```

do this instead

```{r}
weather_df |> 
  group_by(name) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE),
    q95_prcp = quantile(prcp, 0.95, na.rm = TRUE)
  ) |> 
  knitr::kable(digits = 2)
```

knitr::kable gives table

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) |> 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) |> 
  knitr::kable(digits = 2)
```

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) |> 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() + 
  geom_line()
```

```{r}
weather_df |> 
  mutate(center_tmax = tmax - mean(tmax, na.rm = TRUE)) |> 
  ggplot(aes(x = date, y = center_tmax, color = name)) +
  geom_point()
```

## what about "window" functions

try to rank things

```{r}
weather_df |> 
  group_by(name, month) |> 
  mutate(temp_rank = min_rank(desc(tmax))) |> 
  filter(temp_rank < 2)
```

min_rank ranks from smallest value

## lead and lag

```{r}
weather_df |> 
  group_by(name) |> 
  mutate(
    lagged_tmax = lag(tmax, 3),
    lead_tmax = lead(tmax)
  ) 
```

* lag takes value previous row and pastes it to next column
* the first lag column is always NA
* lead does the opposite

```{r}
weather_df |> 
  group_by(name) |> 
  mutate(
    temp_change = tmax - lag(tmax)
  ) |> 
  summarize(
    mean_temp_change = mean(temp_change, na.rm = TRUE),
    sd_temp_change = sd(temp_change, na.rm = TRUE)
  )
```

## revisit some examples

```{r}
pulse_df =
  haven::read_sas("public_pulse_data.sas7bdat") |> 
  janitor::clean_names() |> 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) |> 
  mutate(
    visit = replace(visit, visit == "bl", "00m")
  )
```

* before tidying, check to see if you need to pivot (run up to janitor::clean_names)
* names_to = what you want to make the variable
* names_prefix = the similarities of the variable by a prefix, what you want to change to names_to
* values_to = name of variable value
* replace is for matching prefix/suffix with other variables

```{r}
pulse_df |> 
  group_by(visit) |> 
  summarize(
    BDI_mean = mean(bdi,na.rm = TRUE),
    BDI_median = median(bdi, na.rm = TRUE)
  ) |> 
  knitr::kable(digits = 2)
```

```{r}
pups_df =
  read_csv("FAS_pups.csv", skip = 3, na = c("", ".", "NA")) |> 
  janitor::clean_names() |> 
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litters_df =
  read_csv("FAS_litters.csv", na = c("", ".", "NA")) |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "day_of_tx"), 3)

fas_df =
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |> 
  group_by(dose, day_of_tx) |> 
  drop_na(dose) |> 
  summarize(
    mean_pivot = mean(pd_pivot, na.rm = TRUE)) |> 
  pivot_wider(
    names_from = dose, 
    values_from = mean_pivot) |> 
  knitr::kable(digits = 2)
```

